*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-21 Code/Manuscript_O2.Rmd
MIT9313 growth response results.

The growth rates of *P. marinus* MED4, Clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), when grown under 250 µM O~2~. No growth was
observed under any imposed conditions under a 4 h photoperiod. MED4 was
able to grow under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ but only when the blue light photoperiod was longer than 4 h; the
white LED light photoperiod was longer than 8 h; or when the red light
photoperiod was longer than 16 h. Growth rate increased
with increasing photoperiods for all spectral wavebands when exposed to
PAR of 90 and 180 µmol photons m^-2^ s^-1^. MED4 grew fastest for each
spectral waveband under 16 h photoperiods. The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue
light and 16 h photoperiod.

@@

The growth rates of *P. marinus* SS120, Clade LLIII, increased with lower imposed PAR and with longer photoperiods, when
grown under 250 µM O~2~ (Figure \@ref(fig:SS120PAR)). SS120 was able to
grow under the lowest PAR treatment of 30 µmol photons m^-2^ s^-1^ but
only when the photoperiod was longer than 4 h under white and red
light. No growth was observed under any blue light photoperiods when
exposed to PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate increased
with increasing photoperiods for white and red light when exposed
to PAR of 90 µmol photons m^-2^ s^-1^; however, no growth was observed
when the red light photoperiod was 16 h. Experiments conducted under PAR
of 180 µmol photons m^-2^ s^-1^ showed no growth under blue nor
red lights. Growth rate decreased with longer photoperiods and showed
growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol
photons m^-2^ s^-1^  white LED light.  The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

The growth rates of *P. marinus* MIT9313, Clade LLIV, increased with low to moderate imposed PAR and longer
photoperiods, when grown under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)). Growth rate increased with increasing
photoperiods under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ for blue, white LED and red light. The green light treatment
showed growth only when the photoperiod was longer than 12 h and
decreased by \~ 50% under a 16 h photoperiod. Under PAR of 90 µmol
photons m^-2^ s^-1^, growth rate increased with longer photoperiods for
white LED, red, and green light treatments. Growth rates decreased from
4 to 8 h blue light photoperiod and exhibited no growth under
photoperiods of 12 and 16 h. Blue light did not induce growth at 180
µmol photons m^-2^ s^-1^ under any photoperiod; white LED and red light
treatments only exhibited growth under the 8 h photoperiod and growth
rates increased with longer photoperiods under green light. The µ~max~
was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 30 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

Similar to the growth trends of the 250 µM O~2~ and PAR of 30 µmol
photons m^-2^ s^-1^ experiments, MIT9313 maintained at 25 µM O~2~,
growth rate increased with increasing photoperiods for all spectral
wavebands tested (Figure \@ref(fig:MIT9313PAR)). The 25 µM O~2~ growth
experiments showed generally similar growth rate as the 250 µM O~2~ and
PAR of 30 µmol photons m^-2^ s^-1^, with the interesting exception that
the growth rate increased by 50% under 25 µM O~2~ and 16 h blue light
photoperiod. In contrast to the blue light growth responses under 250 µM
O~2~ and PAR of 90 µmol photons m^-2^ s^-1^, growth rate increased with
longer photoperiods and even grew under 12 and 16 h photoperiods. In
fact, the fastest growth rate of all the O~2~ experiments was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 16 h blue light
photoperiod and PAR of 90 µmol photons m^-2^ s^-1^ under 25 µM O~2~.
Growth rate increased by 50% under the 4 h blue light photoperiod; by
400% under the 8 h blue light photoperiod; by 40% under the 12 h red
light photoperiod and by 80% under the 16 h white LED photoperiod,
compared to the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^
growth experiments. In contrast to the 250 µM O~2~ growth experiments,
MIT9313 was able to grow when exposed to PAR of 180 µmol photons m^-2^
s^-1^ and blue light under all photoperiods except 16 h; additionally,
white LED and red light treatments induced growth across all tested
photoperiods under 25 µM O~2~.

The 25 µM O~2~, 4 h photoperiod experiments under white LED light and
all green light were not performed due to time constraints.

@@

The growth experiments under 2.5 µM O~2~ for MIT9313 showed growth
responses for all spectral wavebands except white LED light, albeit
slower than under the 25 µM O~2~ experiments (Figure
\@ref(fig:MIT9313PAR)). MIT9313 was able to grow under PAR of 30 µmol
photons m^-2^ s^-1^ but only under an 8 h blue and red light
photoperiod. Growth rate increased with longer photoperiods under blue
light treatments and PAR of 90 µmol photons m^-2^ s^-1^ however, growth
was inhibited under a 16 h blue light photoperiod and only grew under an
8 h photoperiod under red and green light treatments. The growth rate
for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light
treatment decreased with increasing photoperiods with growth inhibition
under a 16 h photoperiod. The red light treatments under PAR of 180 µmol
photons m^-2^ s^-1^ showed similar growth rate for 8 and 12 h
photoperiods and no growth were observed for green light treatments. The
µ~max~ was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 4, Par_ue == 180) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 4 h blue light
photoperiod and PAR of 180 µmol photons m^-2^ s^-1^.

>>>>

The growth rates of *P. marinus* MED4, Clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), under 250 µM O~2~. No growth was
observed under any imposed conditions under a 4 h photoperiod. MED4 was
able to grow under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ but only when the blue light photoperiod was longer than 4 h; the
white LED light photoperiod was longer than 8 h; or when the red light
photoperiod was longer than 16 h. Growth rate increased
with increasing photoperiods for all spectral wavebands when exposed to
PAR of 90 and 180 µmol photons m^-2^ s^-1^. MED4 grew fastest for each
spectral waveband under 16 h photoperiods. The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue
light and 16 h photoperiod.

@@

The growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and 250 µM O~2~ (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when
exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate increased
with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^; however, no growth was observed
when the red light photoperiod was 16 h. Experiments under peak PAR
of 180 µmol photons m^-2^ s^-1^ showed no growth under blue nor
red lights. Growth rate decreased with longer photoperiods and showed
growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol
photons m^-2^ s^-1^  white LED light.  The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

The growth rates of *P. marinus* MIT9313, Clade LLIV increased with longer
photoperiods under low 30 µmol photons m^-2^ s^-1^ peak PAR, under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiod. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while cells shows only marginal growth under white LED and red light under the 8 h photoperiod . The µ~max~
was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 30 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

For MIT9313 maintained at 25 µM O~2~ growth rate increased with increasing photoperiods for all spectral wavebands tested (Figure \@ref(fig:MIT9313PAR)). Under peak PAR of 90 µmol photons m^-2^ s^-1^ growth rate increased with
longer photoperiods, with the fastest overall growth rate for MIT9313 
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 16 h blue light
photoperiod. In contrast to the 250 µM O~2~ growth experiments,
MIT9313 grew when exposed to peak PAR of 180 µmol photons m^-2^
s^-1^ and blue light under all photoperiods except 16 h; additionally,
white LED and red light treatments induced growth across all tested
photoperiods under 25 µM O~2~. The 25 µM O~2~, 4 h photoperiod experiments under white LED light and were not performed due to time constraints.

@@

Under 2.5 µM O~2~  MIT9313 showed growth, particularly under blue LED light, albeit
generally slower than under the parallel 25 µM O~2~ experiments (Figure
\@ref(fig:MIT9313PAR)). Growth estimates did, however, show scatter among replicates, suggesting the 2.5 µM O~2~  was near the tolerance limit for growth of  MIT9313. Growth rates increased with longer photoperiods under blue
light treatments under peak PAR of 90 µmol photons m^-2^ s^-1^ . Growth 
for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light
treatment decreased with increasing photoperiods with full growth inhibition
under a 16 h photoperiod. The red light peak PAR of 180 µmol
photons m^-2^ s^-1^ showed similar growth rates for 8 and 12 h
photoperiods. The µ~max~ was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 4, Par_ue == 180) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 4 h blue light
photoperiod and PAR of 180 µmol photons m^-2^ s^-1^.  XXXXTHIS DOES NOT LOOK correct; 12 h, 90 uE photons blue light looks higher.
